Large-scale hybrid ad hoc network for mobile platforms: Challenges and Experiences

### 2019-02-11 paper.tex
Finalized paper structure

WiFi Direct (WiDi) is a clever combination of Hotspot with WiFi DNS
wherein peers randomly choose to be a Hotspot or a client, discover
WiFi credentials over WiFi DNS discovery and make WiFi connection to
Hotspot without needing user permission.  However, both Hotspot and
WiDi suffer from higher battery consumption than DNS and hence were
not deployed. All Android protocols leverage foreground services to
continue discovery and advertisement operations in the background
indefinitely. Deployed protocols (WiFi DNS, BT Nearby, and BTLE) are
described in detail in the following.

\noindent\textbf{WiFi DNS} DNS service discovery is a widely supported
protocol \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises
on this standard such that the exchange of information happens without
making network connections at all. This is achieved by splitting the
message payloads into small enough chunks and stuffing them into
service advertisements as txt records and broadcast over multiple
advertisements in a quick succession. Since there are no connections
being made, the messages can be exchanged without prompting the user
to grant permissions for each connection. Since no peer connections
are needed and advertisements and discovery are broadcast, WiFi DNS is
a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
discovery but is power-hungry and does not interoperate with WiFi DNS
on Android. Algorithm~\ref{algo:dns} outlines the peer process for
WiFi DNS.

\begin{algorithm}[h]
\label{algo:dns}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
\SetAlgoLined
\KwIn{Peer id $p_{id}$, advert/discovery intervals $T_{ad}$, $T_{disc}$}
 $S_{name} \gets$ \textsf{'MNA'}\;
 $\mathbb{M} \gets \varnothing$\;
 \ForEach{timer expiry $T_{ad}$}{
   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
   dnsAdvertise ($S_{name}$, $B$)\;
 }
 \ForEach{timer expiry $T_{disc}$}{
   dnsDiscover ($S_{name}$)\;
 }
 \ForEach{discovered service callback $S_{disc}$}{
   $B \gets S_{disc}.B$\;
   $M_{part} \gets msg(B)$\;
   $M \gets M + M_{part}$\; 
   \If{$M_{part}$ was the last chunk of $M$}{
     \If{$M$ is unexpired $\wedge$ $M$ has valid signature}{
       $\mathbb{M} \gets \mathbb{M} \cup M$\;
     }
   }
 }
 \tcc{For testing only}
 \ForEach{app request to send $M$}{
   \If{$M$ has not expired}{
     $\mathbb{M} \gets \mathbb{M} \cup M$\;
   }
 }
 \caption{WiFi DNS peer algorithm}
\end{algorithm}

\noindent\textbf{BTLE} The only protocol that stays reliably active in
background without violating iOS developer guidelines is BTLE. Unlike
Android, iOS applications do not need to request discovery and
advertisements contunually. Once a BTLE Peripheral (role that has
information) is advertised and a Central (role that receives
information) requests discovery, the app is allowed to go to
sleep. When a matching central or a peripheral device is found, iOS
wakes up the application app to handle such an event, even when the
application is in the background. MNA builds on this and leverages
acynchronous dispatch queues to turn each peer into a Central as well
as a Peripheral simultaneously. This architecture allows bidirectional
data transfers without concurrency issues.
%
\subsection{Power constraints}
\label{sec:power}
%
Due to indefinite forground services and continuous discovery and
advertisement, this is primarily an issue on Android. Our approach
here is two-pronged. Firstly, via extensive experiments on a large
number of devices, we fine-tune the algorithms governing the intervals
at which discovery and advertisements occur. In a nutshell, receiving
new information during a period causes MNA to be more aggressive in
discovery and advertisement. Similarly, lack of new information for a
period makes the device less aggressive. Secondly, we allow ``wake
up'' messages to be broadcast in the network ahead of an anticipated
severe weather event. When devices receive such messages, they
schedule themselves to remain aggressive during the specified window
of time. Outside of this window, the device can afford to have long
sleep cycles and conserve power. With these techniques, our testing
shows less than 2\% battery consumption per hour on most device
models.
%
\subsection{Testing p2p networks}
\label{sec:texting}
%
We developed test automation tools and processes such that multiple
devices can be controlled from a single test station and follow
prescribed steps to generate, send, and receives messages to play out
a test scenario. Finally, the framework allows automated analysis of
the observations to determine the test result. This capability was
instrumental in uncovering bugs at a fast pace to meet the aggressive
timeline and avoid the cost of acquisition. Further, the automation
framework is general and can be expressly applied to test other apps
in this fashion.

@@

%
\subsection{Trust in information}
\label{sec:trust}
%
In our approach, since the origin of weather information is the
Weather Channel service, digital signatures can be attached to all
weather alerts broadcast from the service. The mobile application is
distributed with the corresponding public key so that digital
signatures from the service can be verified. If a message fails such a
verification, it is discarded and not forwarded any further.

%
\section{Experimental Results}
\label{sec:eval}
%

>>>>

WiFi Direct (WiDi) is a clever combination of Hotspot with WiFi DNS
wherein peers randomly choose to be a Hotspot or a client, discover
WiFi credentials over WiFi DNS discovery and make WiFi connection to
Hotspot without needing user permission.  However, both Hotspot and
WiDi suffer from higher battery consumption than DNS and hence were
not deployed. All Android protocols leverage foreground services to
continue discovery and advertisement operations in the background
indefinitely. Algorithms for WiFi DNS and BTLE are described in detail
in Section~\ref{sec:algo}.

%
\subsection{Power constraints}
\label{sec:power}
%
Due to indefinite forground services and continuous discovery and
advertisement, this is primarily an issue on Android. Our approach
here is two-pronged. Firstly, via extensive experiments on a large
number of devices, we fine-tune the algorithms governing the intervals
at which discovery and advertisements occur. In a nutshell, receiving
new information during a period causes MNA to be more aggressive in
discovery and advertisement. Similarly, lack of new information for a
period makes the device less aggressive. Secondly, we allow ``wake
up'' messages to be broadcast in the network ahead of an anticipated
severe weather event. When devices receive such messages, they
schedule themselves to remain aggressive during the specified window
of time. Outside of this window, the device can afford to have long
sleep cycles and conserve power. With these techniques, our testing
shows less than 2\% battery consumption per hour on most device
models.
%
\subsection{Testing p2p networks}
\label{sec:texting}
%
We developed test automation tools and processes such that multiple
devices can be controlled from a single test station and follow
prescribed steps to generate, send, and receives messages to play out
a test scenario. Finally, the framework allows automated analysis of
the observations to determine the test result. This capability was
instrumental in uncovering bugs at a fast pace to meet the aggressive
timeline and avoid the cost of acquisition. Further, the automation
framework is general and can be expressly applied to test other apps
in this fashion.

@@

%
\subsection{Trust in information}
\label{sec:trust}
%
In our approach, since the origin of weather information is the
Weather Channel service, digital signatures can be attached to all
weather alerts broadcast from the service. The mobile application is
distributed with the corresponding public key so that digital
signatures from the service can be verified. If a message fails such a
verification, it is discarded and not forwarded any further.
%
\section{Algorithms}
\label{sec:algo}
%
\subsection{WiFi DNS}
\label{sec:wifi}
%
DNS service discovery is a widely supported
protocol \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises
on this standard such that the exchange of information happens without
making network connections at all. This is achieved by splitting the
message payloads into small enough chunks and stuffing them into
service advertisements as txt records and broadcast over multiple
advertisements in a quick succession. Since there are no connections
being made, the messages can be exchanged without prompting the user
to grant permissions for each connection. Since no peer connections
are needed and advertisements and discovery are broadcast, WiFi DNS is
a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
discovery but is power-hungry and does not interoperate with WiFi DNS
on Android. Algorithm~\ref{algo:dns} outlines the peer process for
WiFi DNS.

\begin{algorithm}[htb!]
\label{algo:dns}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
\SetAlgoLined
\KwIn{Peer id $p_{id}$}
\KwIn{App id $a_{id}$, max advert/discovery intervals $max_{ad}$, $max_{disc}$}
\KwIn{min advert/discovery intervals $min_{ad}$, $min_{disc}$}
 $S_{name} \gets$ \textsf{'MNA'}\;
 $\mathbb{M} \gets \varnothing$\;
 $T_{ad} \gets min_{ad}$\;
 $T_{disc} \gets min_{disc}$\;
 $New \gets 0$
 \ForEach{timer expiry $T_{ad}$}{
   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
   dnsAdvertise ($S_{name}$, $B$)\;
   \If{$New = 0$}{
     $T_{ad} \gets min(T_{ad} \times 2, max_{ad})$\;
   }
 }
 \ForEach{timer expiry $T_{disc}$}{
   dnsDiscover ($S_{name}$)\;
   \If{$New = 0$}{
     $T_{disc} \gets min(T_{disc} \times 2, max_{disc})$\;
   }
   $New \gets 0$\;
 }
 \ForEach{discovered service callback $S_{disc}$}{
   $B \gets S_{disc}.B$\;
   $M_{part} \gets msg(B)$\;
   $M \gets M + M_{part}$\; 
   \If{$M_{part}$ was the last chunk of $M$}{
     \If{$M$ is not a duplicate}{
       \If{$M$ is unexpired $\wedge$ $M$ has valid signature}{
         $\mathbb{M} \gets \mathbb{M} \cup M$\;
         $T_{ad} \gets max(\frac{T_{ad}}{2}, min_{ad})$\;
         $T_{disc} \gets max(\frac{T_{disc}}{2}, min_{disc})$\;
         $New \gets 1$\;
         \If{$M.app = a_{id}$}{
           Notify app with $M$\;
         }
       }
     }
   }
 }
 \tcc{For testing only}
 \ForEach{app request to send $M$}{
   \If{$M$ has not expired}{
     $\mathbb{M} \gets \mathbb{M} \cup M$\;
   }
 }
 \caption{WiFi DNS peer algorithm}
\end{algorithm}
%
\subsection{BTLE}
\label{sec:btle}
%
The only protocol that stays reliably active in
background without violating iOS developer guidelines is BTLE. Unlike
Android, iOS applications do not need to request discovery and
advertisements contunually. Once a BTLE Peripheral (role that has
information) is advertised and a Central (role that receives
information) requests discovery, the app is allowed to go to
sleep. When a matching central or a peripheral device is found, iOS
wakes up the application app to handle such an event, even when the
application is in the background. MNA builds on this and leverages
acynchronous dispatch queues to turn each peer into a Central as well
as a Peripheral simultaneously. This architecture allows bidirectional
data transfers without concurrency issues.

\section{Experimental Results}
\label{sec:eval}
%
